## Expression analysis of mouse pcRNAs

```{r set-options, echo=FALSE, cache=FALSE}
        options(replace.assign=TRUE,width=70)
        opts_chunk$set(dev=c("png","pdf"), dpi=300, fig.path='mouseExp-fig/', cache.path='cache/mousExp-', par=TRUE, echo=FALSE, warning=F, message=F, cache=FALSE, results='hide', error=F)
```

```{r load-libs, cache=F}
library("ggplot2")
library("reshape")
library("gplots")
library("RColorBrewer")
hmcol = colorRampPalette(brewer.pal(9, "GnBu"))(100)

load("report/data/pc_human_expAnalysis.Rdata")
load("report/data/hsa_nc_exp.Rdata")
load("report/data/mmu_nc_exp.Rdata")
load("report/data/mouse2human.Rdata")

source("functions.R")

```

```{r mouse-exp-analysis}
##########################
#       Mouse           ##
##########################
colnames(mmu_nc_exp) <- gsub("_FPKM", "", colnames(mmu_nc_exp))
colnames(hsa_nc_exp) <- gsub("_FPKM", "", colnames(hsa_nc_exp))


mmu_nc_exp_mean <- mmu_nc_exp
thr <- 0
mmu_nc_exp_mean_filt <- mmu_nc_exp_mean[apply(mmu_nc_exp_mean,1,max)>thr,]
mmu_nc_exp_mean_melt <- melt(data.frame(Name=row.names(mmu_nc_exp_mean_filt),mmu_nc_exp_mean_filt))
colnames(mmu_nc_exp_mean_melt) <- c("Name", "Tissue", "FPKM")

hsa_nc_exp_mean <- hsa_nc_exp
thr_hsa <- 0
hsa_nc_exp_mean_filt <- hsa_nc_exp_mean[apply(hsa_nc_exp_mean,1,max)>thr_hsa,]


mmu_meanAndMax <- t(apply(mmu_nc_exp,1,function(x) c(mean(x),max(x))))
colnames(mmu_meanAndMax) <- c("Mean", "Max")
mmu_meanAndMax <- mmu_meanAndMax[mmu_meanAndMax[,2]>thr,]
mmu_meanAndMax <- melt(mmu_meanAndMax)
colnames(mmu_meanAndMax) <- c("Name", "Type", "FPKM")

# Calculate tissue specificity
mmu_tissue_specificity <- as.data.frame(t(apply(mmu_nc_exp_mean_filt, 1, tissueScore)), stringsAsFactors=F)
colnames(mmu_tissue_specificity) <- c("maxTissue", "Score", "maxFPKM")
mmu_tissue_specificity$Score <- as.numeric(mmu_tissue_specificity$Score)
mmu_tissue_specificity$maxFPKM <- as.numeric(mmu_tissue_specificity$maxFPKM)
mmu_tissue_specificity$maxTissue <- as.factor(mmu_tissue_specificity$maxTissue)
mmu_tissue_specificity <- mmu_tissue_specificity[,c("Score", "maxTissue", "maxFPKM")]

score_mmu <- val2col(mmu_tissue_specificity[row.names(mmu_nc_exp_mean_filt), "Score"], col=colorRampPalette(brewer.pal(9, "YlOrRd"))(100))

# Set tissue colors for row side colors of the heatmap
max_tis_mmu <- mmu_tissue_specificity[row.names(mmu_nc_exp_mean_filt), "maxTissue"]
tissues_mmu <- levels(max_tis_mmu)
levels(max_tis_mmu) <- colorRampPalette(brewer.pal(9, "Set1"))(length(levels(max_tis_mmu)))

# Correlation of expression between mouse and human
tmp_m2h<- merge(mouse2human, mmu_nc_exp_mean_filt[,c("C2C12", "CH12", "MEL", "ESpA", "br", "cb", "ht", "kd", "lv", "ts")], by.x="mmu-pcTCONSid", by.y=0)

mAndH<- merge(tmp_m2h, hsa_nc_exp_mean_filt[c("Hsmm", "GM12878", "K562", "EScepA", "br", "cb", "ht", "kd", "lv", "ts")], by.x="Human", by.y=0, all.x=T)

m_matrix <- as.matrix(mAndH[,4:13])
row.names(m_matrix) <- mAndH[,"mmu-pcTCONSid"]

h_matrix <- as.matrix(mAndH[,14:ncol(mAndH)])
row.names(h_matrix) <- mAndH[,"Human"]

mh_rho <- data.frame(Human=row.names(h_matrix), mh_rho=diag(cor(t(h_matrix),t(m_matrix), method="spearman")))

mh_rho_cor <- cor(t(h_matrix),t(m_matrix), method="spearman")
mh_rho_perm <- replicate(1000000, mean(sample(mh_rho_cor, ncol(mh_rho_cor), replace=T), na.rm=T))
mh_rho_pval <- mean(mh_rho_perm>=mean(diag(mh_rho_cor), na.rm=T))
mh_rho_cor_df <- data.frame(Background_rho=as.numeric(mh_rho_cor))
mh_r <- data.frame(Human=row.names(h_matrix), mh_r=diag(cor(t(h_matrix),t(m_matrix))))

pc_mh <- merge(pc, mh_rho, by.x="pcTCONSid", by.y="Human")
pc_mh <- merge(pc_mh, mh_r, by.x="pcTCONSid", by.y="Human")

mmu_tissue_specificity_fullInfo <- merge(mouse2human, mmu_tissue_specificity, by.x="mmu-pcTCONSid", by.y=0)
write.table(mmu_tissue_specificity_fullInfo, file="report/data/mmu_tissue_specificity.txt", sep="\t", quote=F, row.names=F)
```

### General expression analysis
```{r meanMaxmmu, fig.cap="Distribution of the Mean and Max FPKMs (log2) of each mouse pcRNA across tissues. pcRNAs with Max FPKM <thr have been excluded from this plot."}
ggplot(mmu_meanAndMax, aes(x=Type, y=log2(FPKM))) + geom_boxplot(alpha=0.4) + ggtitle("FPKM distribution for mouse pcRNAs") + ylab("log2 FPKM")
```

```{r expression-density-mmu, fig.cap="Distribution of mouse pcRNA expression (log2FPKM) across tissues."}
ggplot(mmu_nc_exp_mean_melt[mmu_nc_exp_mean_melt$FPKM>thr,], aes(log2(FPKM), fill=Tissue)) + geom_density(alpha=0.3) + facet_wrap(~Tissue)
```
```{r expression-density-mmu2, fig.cap="Distribution of mouse pcRNA expression (log2FPKM) across tissues. The minimum FPKM!=0 was added to pcRNAs with FPKM=0 in order to plot them on a log scale."}
min_FPKM <- min(mmu_nc_exp_mean_melt[mmu_nc_exp_mean_melt$FPKM!=0, "FPKM"])
ggplot(mmu_nc_exp_mean_melt, aes(log2(FPKM+min_FPKM), fill=Tissue)) + geom_density(alpha=0.3) + facet_wrap(~Tissue)
```

```{r expression-box-mmu, fig.cap="Distribution of mouse pcRNA expression (log2FPKM) across tissues."}
ggplot(mmu_nc_exp_mean_melt[mmu_nc_exp_mean_melt$FPKM>thr,], aes(y=log2(FPKM), x=Tissue,fill=Tissue)) + geom_boxplot(alpha=0.3) + theme(axis.text.x = element_blank())
```

```{r expression-box-mmu2, fig.cap="Distribution of mouse pcRNA expression (log2FPKM) across tissues. The minimum FPKM!=0 was added to pcRNAs with FPKM=0 in order to plot them on a log scale."}
ggplot(mmu_nc_exp_mean_melt, aes(y=log2(FPKM+min_FPKM), x=Tissue,fill=Tissue)) + geom_boxplot(alpha=0.3) + theme(axis.text.x = element_blank())
```

```{r NC-heatmaps-mmu, fig.cap="Heatmap showing log2 FPKM values for mouse pcRNAs across tissues.  The minimum FPKM!=0 was added to pcRNAs with FPKM=0 in order to plot them on a log scale. pcRNAs with 0 FPKMs in all samples were removed"}
heatmap.2(log2(as.matrix(mmu_nc_exp_mean_filt+min_FPKM)), scale="none", trace="none", density="none", Rowv=T, Colv=T, margins=c(10,10), col=hmcol, labRow=NA)
```

```{r NC-heatmaps-mmu-zscore, fig.cap="Heatmap showing z-scores FPKM values for mouse pcRNAs across tissues. pcRNAs with 0 FPKMs in all samples were removed"}
heatmap.2(as.matrix(mmu_nc_exp_mean_filt), labCol=colnames(mmu_nc_exp_mean), scale="row", trace="none", density="none", Rowv=T, Colv=T, margins=c(10,10), col=hmcol, labRow=NA)
```

### Tissue specificity and correlation of expression

```{r NC-pearson-tissue-mmu, fig.cap="Heatmap showing pair-wise Pearson correlation coefficient between pcRNA expression across tissues. pcRNAs with 0 FPKMs in all samples were removed"}
heatmap.2(cor(mmu_nc_exp_mean_filt), scale="none", dendrogram="column", trace="none", density="none", labRow=colnames(mmu_nc_exp_mean), labCol=colnames(mmu_nc_exp_mean), margins=c(10,10), col=hmcol)
```

```{r NC-spearman-tissue-mmu, fig.cap="Heatmap showing pair-wise Spearman correlation coefficient between tissues. pcRNAs with 0 FPKMs in all samples were removed"}
heatmap.2(cor(mmu_nc_exp_mean_filt, method="spearman"), scale="none", dendrogram="column", trace="none", density="none", labRow=colnames(mmu_nc_exp_mean), labCol=colnames(mmu_nc_exp_mean), margins=c(10,10), col=hmcol)
```

```{r NC-pearson-pc-mmu, fig.cap="Heatmap showing pair-wise Pearson correlation coefficient between tissues. The yellow-to-red bar at the top indicates the tissue specificty score of each pcRNAs, with yellow being 0 and red being 1. pcRNAs with no expression in any tissue were removed", global.device=TRUE }
heatmap.2(cor(t(mmu_nc_exp_mean_filt)), scale="none", dendrogram="column", trace="none", density="none", labRow="", labCol="", margins=c(2,2), col=hmcol, ColSideColors=score_mmu, RowSideColors=as.character(max_tis_mmu))
legend("bottomleft",legend=as.character(tissues_mmu), fill=as.character(levels(max_tis_mmu)), border=FALSE, bty="n", y.intersp = 0.9, cex=0.7, title="Tissues")
```

```{r NC-dist-pc-mmu, fig.cap="Heatmap showing pair-wise Euclidean distance between tissues. pcRNAs with no expression in any tissue were removed", global.device=TRUE }
eu_dist <- t(scale(t(mmu_nc_exp_mean_filt)))
heatmap.2(as.matrix(dist(eu_dist)), scale="none", dendrogram="column", trace="none", density="none", labRow="", labCol="", margins=c(2,2), col=rev(hmcol), RowSideColors=as.character(max_tis_mmu))
legend("bottomleft",legend=as.character(tissues_mmu), fill=as.character(levels(max_tis_mmu)), border=FALSE, bty="n", y.intersp = 0.9, cex=0.7, title="Tissues")
```

```{r NC-spearman-pc-mmu, fig.cap="Heatmap showing pair-wise Spearman correlation coefficient between pcRNA. The yellow-to-red bar at the top indicates the tissue specificty score of each pcRNAs, with yellow being 0 and red being 1. pcRNAs with no expression in any tissue were removed", global.device=TRUE }
heatmap.2(cor(t(mmu_nc_exp_mean_filt), method="spearman"), scale="none", dendrogram="column", trace="none", density="none", labRow="", labCol="", margins=c(2,2), col=hmcol, ColSideColors=score_mmu, RowSideColors=as.character(max_tis_mmu))
legend("bottomleft",legend=as.character(tissues_mmu), fill=as.character(levels(max_tis_mmu)), border=FALSE, bty="n", y.intersp = 0.9, cex=0.7, title="Tissues")
```


```{r tiss-spec-mmu, fig.cap="Tissue specificity score of each mouse pcRNA against its max FPKM across tissues"}
ggplot(mmu_tissue_specificity, aes(y=Score, x=maxFPKM)) + geom_point() + scale_x_log10() + ggtitle("Tissue specificity score")
```

```{r tiss-spec-byTiss-mmu, fig.cap="Tissue specificity score of each mouse pcRNA against its max FPKM across tissues, by tissues"}
ggplot(mmu_tissue_specificity[!is.na(mmu_tissue_specificity$maxTissue), ], aes(y=Score, x=maxFPKM, colour=maxTissue)) + geom_point() + scale_x_log10() + facet_wrap("maxTissue") + ggtitle("Tissue specificity score")
```

```{r tiss-spec-byTiss-mmu-dens, fig.cap="Distribution of tissue specificity scores of each mouse pcRNA."}
ggplot(mmu_tissue_specificity[!is.na(mmu_tissue_specificity$maxTissue), ], aes(Score, fill=maxTissue)) + geom_density(alpha=0.5) + facet_wrap(~maxTissue)
```

### Correlation of expression between human and mouse pcRNAs
In all plots in this section human pcRNAs that have two or more corresponding transcripts in mouse are considered multiples times (i.e. as multiple points in the plots).

```{r mh-rho, fig.cap="Distribution of Spearman correlation coefficients between human and mouse pcRNAs."}
ggplot(pc_mh, aes(mh_rho)) + geom_density(fill="black", alpha=0.2) + xlab("Spearman rho") + geom_density(data=mh_rho_cor_df, aes(Background_rho), alpha=0.2, adjust=3)
```
Mean rho: `r mean(pc_mh$mh_rho, na.rm=T)`; Median rho: `r median(pc_mh$mh_rho, na.rm=T)`; p-value `r mh_rho_pval`

```{r mh-r, fig.cap="Distribution of Pearson correlation coefficients between human and mouse pcRNAs."}
ggplot(pc_mh, aes(mh_r)) + geom_density() + xlab("Pearson r")
```

```{r mh-rho-tis, fig.cap="Distribution of Spearman correlation coefficients between human and mouse pcRNAs, divided by tissue with the maximal expression of the human pcRNA."}
ggplot(pc_mh[!is.na(pc_mh$maxTisPC),], aes(mh_rho, fill=maxTisPC)) + geom_density(alpha=0.6) + xlab("Spearman rho") + facet_wrap(~maxTisPC)
```

```{r mh-r-tis, fig.cap="Distribution of Pearson correlation coefficients between human and mouse pcRNAs, divided by tissue with the maximal expression of the human pcRNA."}
ggplot(pc_mh[!is.na(pc_mh$maxTisPC),], aes(mh_r, fill=maxTisPC)) + geom_density(alpha=0.6) + xlab("Pearson r") + facet_wrap(~maxTisPC)
```

```{r mh-rho-cat, fig.cap="Distribution of Spearman correlation coefficients between human and mouse pcRNAs divided by positional context."}
ggplot(pc_mh, aes(y=mh_rho, x=Context)) + geom_boxplot() + ylab("Spearman rho")
```

```{r mh-r-cat, fig.cap="Distribution of Pearson correlation coefficients between human and mouse pcRNAs divided by positional context."}
ggplot(pc_mh, aes(y=mh_r, x=Context)) + geom_boxplot() + ylab("Pearson r")
```

```{r mh-rho-fp, fig.cap="Scatterplot of Spearman correlation coefficients between human and mouse pcRNAs plotted against the maximum FPKM of the human pcRNA"}
ggplot(pc_mh, aes(x=log2(maxFpkmPC), y=mh_rho)) + geom_point() + ylab("Spearman rho")
```

```{r mh-r-fp, fig.cap="Scatterplot of Pearson correlation coefficients between human and mouse pcRNAs plotted against the maximum FPKM of the human pcRNA"}
ggplot(pc_mh, aes(x=log2(maxFpkmPC), y=mh_r)) + geom_point() + ylab("Pearson r")
```

```{r mh-r-spec, fig.cap="Scatterplot of Pearson correlation coefficients between human and mouse pcRNAs plotted against the tissue specificity score of the human pcRNA"}
ggplot(pc_mh, aes(y=tisSpecPC, x=mh_r)) + geom_point() + ylab("Tissue spec score") + xlab("Pearson r") + facet_wrap(~maxTisPC)
```

```{r mh-rho-spec, fig.cap="Scatterplot of Spaerman correlation coefficients between human and mouse pcRNAs plotted against the tissue specificity score of the human pcRNA"}
ggplot(pc_mh, aes(y=tisSpecPC, x=mh_rho)) + geom_point() + ylab("Tissue spec score") + xlab("Spearman rho") + facet_wrap(~maxTisPC)
```



